\sysfull

### 2018-04-16 paper/implementation.tex
partial commit exec revision

Among \sys's novelties, the task-splitting feature is the most effective when it comes to preservation of forward progress. Its implementation adds hardly any computational overhead, and is supported by a hardware timer running ``in the background'', whose energy consumption is extremely low, i,e. $\approx$3\,$\mu$ A/MHz~\cite{msp430datasheet}. The goal is to interrupt a static task if it cannot be run to a completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue. \sys speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget. To ``split'' a static task, \sys schedules a timer interrupt. If this interrupt occurs before the capacitor is depleted, \sys commits all the dirty volatile pages (using the two-phase commit explained above), saves stack and volatile registers to non-volatile memory and then continues with the task. In case a power failure occurs after this operation, but before the static task completes, the execution at the next reboot starts exactly where \sys had saved the partial execution. A task can be split in multiple sub-tasks, as many as needed to progress forward. Regarding the timer interrupt, it is initially scheduled to fire after a large time interval. If power fails before the interrupt fires, \sys reduces the interval. If the partial execution is committed successfully, \sys enlarges the interval. These adjustments follow a binary search technique, and make the interval converge very fast to a value optimal for preserving forward progress and avoiding redundant commits.

For some applications, specific pieces of a task cannot be split and have to be run atomically. For instance, when sampling an analog signal, continuity between samples has to be ensured, otherwise the digital array would represent a wrong, non-continuous input signal. In such cases, the programmer can use two dedicated API guards before and after the atomic piece of code, namely \texttt{COALA\_DISABLE\_PE} and \texttt{COALA\_ENABLE\_PE}, to respectively halt and resume the partial execution timer.

>>>>

Among \sys's novelties, the task-splitting feature is the most effective when it comes to preservation of forward progress. Its implementation adds very little computational overhead, and is supported by a hardware timer running ``in the background'', whose energy consumption is extremely low, i,e. $\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}. The goal is to interrupt a static task if it cannot be run to a completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue.

\textbf{Partial Commit Algorithm.} \sys speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget. To ``split'' a static task, \sys schedules a timer interrupt. If this interrupt occurs before the capacitor is depleted, \sys commits all the dirty volatile pages (using the two-phase commit explained above), saves stack and volatile registers to non-volatile memory and then continues with the task. In case a power failure occurs after this operation, but before the static task completes, the execution at the next reboot starts exactly where \sys had saved the partial execution. A task can be split in multiple sub-tasks, as many as needed to progress forward. The timer interrupt is initially scheduled to fire after a large time interval. If power fails before the interrupt fires, \sys reduces the interval. If the partial execution is committed successfully, \sys enlarges the interval. These adjustments follow a binary search technique, and make the interval converge very fast to a value optimal for preserving forward progress and avoiding redundant commits.

\textbf{Discussion.} For some applications, specific pieces of a task cannot be split and have to be run atomically. For instance, when sampling an analog signal, continuity between samples has to be ensured, otherwise the digital array would represent a wrong, non-continuous input signal. In such cases, the programmer can use two dedicated API guards before and after the atomic piece of code, namely \texttt{COALA\_DISABLE\_PE} and \texttt{COALA\_ENABLE\_PE}, to respectively halt and resume the partial execution timer.
